---
title: "Seattle Hottest Time of Day Seasonally"
author: "Roger Andre"
date: 'Nov 21, 2023'
format: 
  html:
    code-fold: true
    theme: flatly
execute: 
  enabled: true
---

## Does the hottest time of day change?

Previously, I looked at a single hot day in Seattle to see when the maximum temperature occurred.  For a day in early August 2022, the high temp was between 3:00 and 4:00 PM.  Now the question I want to answer is whether the time of day at which the maximum temp occurs changes over the course of the year?

## Data prep

Again, I'll download some weather data I've been capturing and storing on [Github]("https://github.com/lagerratrobe/weather_station/raw/main/Data/station_obs.RDS") and will extract the hourly temperature values for Seattle over the past year.  Then, I'll identify the hottest temp for each day and the time that took place.  Finally, I'll plot the whole mess and we'll see what happens.

```{r}
#| echo: false

library(dplyr, warn.conflicts = FALSE)
library(lubridate, warn.conflicts = FALSE)
library(ggplot2)

df <- readRDS(url("https://github.com/lagerratrobe/weather_station/raw/main/Data/station_obs.RDS"))
df |> filter(stationID == "KWASEATT2743") |>
  mutate(stationID,
         datetime = ymd_hms(obsTimeLocal),
         temp = imperial.temp,
         .keep = "none") -> df
```

```{r}
#| echo: false
df |> group_by(date(datetime)) |> 
  slice_max(temp, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(stationID, 
         datetime, 
         'high_temp' = temp) -> daily_max_temps
```

The data that is going to be plotted looks like this:

```{r}
#| echo: false

head(daily_max_temps) |> as.data.frame()
```

### Plot Results

When I plot this with Date on the X axis and Hour on the Y axis, I see that in general, the max temp is at around 3:00PM.  During days with warmer temperatures this is very consistent, however as days get colder, there appears to be greater variability in when the high temp is achieved. In some cases it is early in the morning or even at night.

```{r}
#| echo: false
daily_max_temps |> 
  ggplot(mapping = aes(x =datetime, y= hour(datetime), color=high_temp)) +
  scale_color_gradient2(low = "darkblue",
                        mid = "lightpink",
                       high = "red",
                       midpoint = 60) +
  scale_y_reverse() +
  geom_jitter(alpha = .65) +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, col='black') +
  labs(x="Date",
       y="Hour")
```

