---
title: "Seattle Hot Day"
author: "Roger Andre"
date: 'Nov 20, 2023'
format: 
  html:
    code-fold: true
    theme: flatly
execute: 
  enabled: true
---

## What's the hottest time of the day?

As part of my ongoing analysis of the weather in Seattle over the past year, I want to see what the hottest part of the day is.  To do that, I'll download weather data and pick one day where the temperature got above 80 deg F and then plot the hourly temperatures.

```{r}
#| echo: false
# Read in the weather data, filter it down to just Seattle and Sequim, select time, temp and precip columns and create proper date and time objects.
library(dplyr, warn.conflicts = FALSE)
library(lubridate, warn.conflicts = FALSE)

df <- readRDS(url("https://github.com/lagerratrobe/weather_station/raw/main/Data/station_obs.RDS"))

df |> filter(stationID %in% c("KWASEATT2743", "KWASEQUI431")) |>
  mutate(stationID,
         datetime = ymd_hms(obsTimeLocal),
         temp = imperial.temp,
         precip = imperial.precipTotal,
         .keep = "none") -> df
```

### Step 1. Find a couple hot days in Seattle

(For our purposes, "hot" will be 80 deg, or warmer.)

```{r}
#| echo: false
df |> 
  group_by(stationID, date(datetime)) |> 
  summarise(max_temp = max(temp), .groups = "keep") |>
  filter(stationID == "KWASEATT2743" &  max_temp >= 80) |> 
  head(n=3) |> as.data.frame()
```
### Step 2. Pull the hourly values for one day and plot them

```{r}
#| echo: false
library(ggplot2)
hot_day <- filter(df,
                  stationID == "KWASEATT2743" & 
                    date(datetime) == '2022-08-06')

ggplot(hot_day, mapping = aes(x = datetime, y=temp)) + 
  geom_line() +
  scale_x_datetime(date_labels = "%H:%M") +
  geom_vline(xintercept=hot_day$datetime[15],color="red") +
  annotate("text", 
           x=hot_day$datetime[16], 
           y=60, 
           label=format(hot_day$datetime[15], "%H:%M"),
           color="red") +
  ggtitle("Aug 6, 2022 Hourly Temps") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Time",
       y = "Temp")

```